```{r}
#Stats
#install.packages("ggplot2")   # only if not installed
#library(ggplot2)
#install.packages("ggrepel")   # only if not installed
#library(ggrepel)
```

# Introduction and Data: 

This project is an analysis of different NFL statistics and how they relate to points scored per game. Originally, we compared 3 offensive and 3 defensive statistics to wins (including points and points allowed), which obviously showed that wins were most correlated to points scored. After compiling our data and presenting it, we realized that this project said nothing about who wins in the NFL regular season. Because of this, we changed our focus to comparing 3 offensive statistics (rushing yards, passing yards, turnovers lost) and 3 defensive statistics (rushing yards allowed, passing yards allowed, takeaways) to points per game and points allowed per game respectively. We will attempt to refute the hypothesis that any of these 6 particular statistics do not significantly contribute to the overall points per game (offense) or points allowed per game (defense) statistic.

```{r}


# Helper function to compute r and r2 label
make_r_label <- function(x, y) {
  r <- cor(x, y)
  r2 <- r^2
  paste0("r = ", round(r, 3),
         "\nr² = ", round(r2, 3))
}

### 1. PA vs PYdsA
ggplot(NFL_Stats, aes(x = PA, y = PYdsA, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Allowed", y = "Passing Yards Allowed", title = "NFL Points Allowed vs Passing Yards Allowed") +
  annotate("text",
           x = max(NFL_Stats$PA),
           y = min(NFL_Stats$PYdsA),
           label = make_r_label(NFL_Stats$PA, NFL_Stats$PYdsA),
           hjust = 1, vjust = 0) +
  theme_minimal()

## 2. PA vs RYdsA
ggplot(NFL_Stats, aes(x = PA, y = RYdsA, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Allowed", y = "Rushing Yards Allowed", title = "NFL Points Allowed vs Rushing Yards Allowed") +
  annotate("text",
           x = max(NFL_Stats$PA),
           y = min(NFL_Stats$RYdsA),
           label = make_r_label(NFL_Stats$PA, NFL_Stats$RYdsA),
           hjust = 1, vjust = 0) +
  theme_minimal()

##PA vs Takeaways
ggplot(NFL_Stats, aes(x = PA, y = `Takeaways(Defense)`, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Allowed", y = "Takeaways", title = "NFL Points Allowed vs Takeaways") +
  annotate("text",
           x = max(NFL_Stats$PA),
           y = min(NFL_Stats$`Takeaways(Defense)`),
           label = make_r_label(NFL_Stats$PA, NFL_Stats$`Takeaways(Defense)`),
           hjust = 1, vjust = 0) +
  theme_minimal()

##Points For vs RY/G
ggplot(NFL_Stats, aes(x = `Pts For`, y = `RY/G`, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Scored", y = "Rushing Yards", title = "NFL Points Scored vs Rushing Yards") +
  annotate("text",
           x = max(NFL_Stats$`Pts For`),
           y = min(NFL_Stats$`RY/G`),
           label = make_r_label(NFL_Stats$`Pts For`, NFL_Stats$`RY/G`),
           hjust = 1, vjust = 0) +
  theme_minimal()

## Pts For vs PY/G
ggplot(NFL_Stats, aes(x = `Pts For`, y = `PY/G`, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Scored", y = "Passing Yards", title = "NFL Points Scored vs Passing Yards") +
  annotate("text",
           x = max(NFL_Stats$`Pts For`),
           y = min(NFL_Stats$`PY/G`),
           label = make_r_label(NFL_Stats$`Pts For`, NFL_Stats$`PY/G`),
           hjust = 1, vjust = 0) +
  theme_minimal()

## Pts For vs TO Lost
ggplot(NFL_Stats, aes(x = `Pts For`, y = `TO Lost (Offense)`, label = '')) +
  geom_point(size = 3) +
  geom_text_repel(size = 3) +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  labs(x = "Points Scored", y = "Turnovers Lost", title = "NFL Points Scored vs Turnovers Lost") +
  annotate("text",
           x = max(NFL_Stats$`Pts For`),
           y = min(NFL_Stats$`TO Lost (Offense)`),
           label = make_r_label(NFL_Stats$`Pts For`, NFL_Stats$`TO Lost (Offense)`),
           hjust = 1, vjust = 0) +
  theme_minimal()
# View coefficient tests
summary(model)$coefficients

```

```{r}
# Optional: nicer p-value formatting
options(scipen = 5)

offense_model <- lm(`Pts For` ~ `PY/G` + `RY/G` + `TO Lost (Offense)`,
                    data =NFL_Stats)

defense_model <- lm(PA ~ PYdsA + RYdsA + `Takeaways(Defense)`,
                    data = NFL_Stats)

offense_coefs <- summary(offense_model)$coefficients
offense_coefs

defense_coefs <- summary(defense_model)$coefficients
defense_coefs
```
